\TITLE

### 2020-12-18 vonLaszewski-covid.tex
add response
update table

\usepackage[table]{xcolor}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{rotating, multirow}

@@

@@

To establish the feasibility of using deep learning, we have, in our
analysis, verified that through deep learning we can recreate the
predictions identified by a model-based approach using empirical
fits. In this comprehensive analysis, we obtain predictions through
deep learning while using data generated from epidemiological models
as discussed in \cite{marsland20-covid-paper}. The time series we used
was 100 days long, and a multi-layered Long Short-Term Memory (LSTM)
recurrent network was used. Our prediction approach differs by
learning not only from the demographics (fixed data for each city) and
time-dependent data but by integrating the population model for the
underlying prediction, as shown in Fig. \ref{fig:magic-1}. Such model
predictive integration capability is important in any application with
multiple time scales. For example, this allows us to integrate
multi-scale time effects into the forecast, which could address the
combination of general forecasts and the next time steps of choice,
such as days, weeks, months, or longer. For this analysis, we used 37
of the 110 cities with reliable empirical fits (that we derived from
an epidemiological model and not deep learning
\cite{marsland20-covid-paper}) for the cases and deaths data up to
April 15, 2020. Our analysis identified a sophisticated single deep
learning time evolution operator that can describe these 37 separate
data sets, and smooth fitted data leads to very accurate deep learning
descriptions via LSTM as depicted in both Figure \ref{fig:magic-1}
that are smaller than 1\% from the model predictions.

@@

\begin{table}[!p]
\caption{Best models for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey.}
\label{tab:top-1}
\bigskip
\centering
\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{RMSE\\Cases} &  
 \makecell{RMSE\\Death} &  
 \makecell[r]{Cummulative \\Error \\Cases} &  
 \makecell[r]{Cummulative \\Error \\Death} &  
 \makecell[r]{Number of \\Days as \\Input} & 
 \makecell{Risk Factor} &  
 \makecell{Place/\\Rank} & \\
\midrule
 0.055443 & 0.067617 & 7.94 & 22.74 & 5 & POP\_DENSITY\_2010 & 0 \\
 0.054795 & 0.068529 & 8.01 & 20.55 & 3 & PHLTH & 1 \\
 0.055199 & 0.067641 & 8.07 & 20.43 & 4 & INSURANCE & 2 \\
 0.055028 & 0.067207 & 8.12 & 20.17 & 4 & PERCENTBLACKS & 3 \\
 0.056061 & 0.067489 & 8.22 & 19.04 & 5 & PERCENTHISPANICS & 4 \\
 0.054621 & 0.068255 & 8.26 & 20.10 & 3 & NHOSP & 5 \\
 0.054557 & 0.068572 & 8.34 & 24.61 & 3 & INSURANCE & 6 \\
 0.054712 & 0.069121 & 8.37 & 21.70 & 3 & CHOLSCREEN & 7 \\
 0.054586 & 0.067636 & 8.40 & 20.50 & 5 & DIABETES & 8 \\
 0.054160 & 0.066906 & 8.45 & 20.16 & 4 & STROKE & 9 \\
 0.054027 & 0.067642 & 8.46 & 21.93 & 5 & CHD & 10 \\
 0.054927 & 0.067652 & 8.51 & 20.42 & 5 & CHECKUP & 11 \\
 0.054146 & 0.067536 & 8.54 & 21.15 & 5 & NBEDS & 12 \\
 0.054795 & 0.067676 & 8.58 & 19.94 & 5 & SVI\_OVERALL & 13 \\
 0.054600 & 0.067565 & 8.61 & 21.86 & 5 & KIDNEY & 14 \\
 0.055089 & 0.068248 & 8.63 & 22.58 & 5 & CASTHMA & 15 \\
 0.054547 & 0.068266 & 8.68 & 20.49 & 3 & BLACK & 16 \\
 0.054685 & 0.068920 & 8.74 & 21.69 & 3 & BPMED & 17 \\
 0.054323 & 0.067149 & 8.75 & 19.54 & 5 & LPA & 18 \\
 0.053716 & 0.066898 & 8.75 & 19.01 & 5 & CSMOKING & 19 \\
 0.054053 & 0.068716 & 8.83 & 20.94 & 3 & ARTHRITIS & 20 \\
 0.055029 & 0.068672 & 8.84 & 22.02 & 3 & POVERTY & 21 \\
 0.054545 & 0.067267 & 8.86 & 21.22 & 4 & ESTBEDS & 22 \\
 0.054864 & 0.068742 & 8.90 & 19.53 & 3 & MHLTH & 23 \\
 0.055079 & 0.069024 & 8.93 & 21.23 & 3 & SENIOR & 24 \\
 0.055589 & 0.068250 & 8.93 & 23.26 & 4 & COPD & 25 \\
 0.054500 & 0.068363 & 8.99 & 19.70 & 3 & CANCER & 26 \\
 0.054397 & 0.068491 & 9.05 & 21.89 & 3 & NBEDS/1000 & 27 \\
 0.054609 & 0.067346 & 9.13 & 21.29 & 4 & BPHIGH & 28 \\
\rowcolor{black!5}  0.054868 & 0.069432 & 9.26 & 25.88 & 3 & NONE & 29 \\
 0.056429 & 0.069345 & 9.28 & 20.98 & 3 & HIGHCHOL & 30 \\
 0.054317 & 0.067638 & 9.32 & 21.00 & 5 & BINGE & 31 \\
 0.054469 & 0.068540 & 9.50 & 20.65 & 3 & OBESITY & 32 \\
 0.055094 & 0.067443 & 9.54 & 21.53 & 4 & SVI\_MINORITY & 33 \\
\bottomrule
\end{tabular}
}
\end{table}

@@

\begin{figure}[!h]
    \begin{minipage}{.45\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{images/predict/Length_5-fields_pop_density_2010-Cumulative-Cases.png}
    \vspace{-1cm}
    \caption{Model prediction error for cumulative cases when
      including one risk factor - POP\_DENSITY\_2010 in the features,
      in addition to the past number of cases and deaths.}
    \label{fig:error-case-popdensity}
    \end{minipage}
    \begin{minipage}{.45\textwidth}
    \ \
    \centering
    \includegraphics[width=1.0\textwidth]{images/predict/Length_5-fields_pop_density_2010-Cumulative-Death.png}
    \vspace{-1cm}
    \caption{Model prediction error for cumulative deaths when
      including one risk factor - POP\_DENSITY\_2010 in the features,
      in addition to the past number of cases and deaths.}
    \label{fig:error-death-popdensity}
    \end{minipage}
\end{figure}

@@

In our next comprehensive analysis, we try to answer Question
\ref{q:4} if all risk factors impact the analysis. Thus we feed the
covariates into the recurrent network as additional features. In
addition to cases and deaths, this leads to 35 variables, including 33
unique covariates. We looked at two related ways of doing this:

\item [(a)] Have covariates fixed in time (static)

\item[(b)] Multiply the covariates by the times series (we used the
  average of cases and deaths)

@@

\begin{table}[!h]
\caption{Correlation Matrix}
\bigskip
\label{tab:corr-matrix}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{lrrrrrr}
\toprule
Risk Factor		& Past Case & Now Case & Future Case & Past Death & Now Death & Future Death \\
\midrule
PVI                     & -0.425      & -0.292      & -0.124      & -0.434      & -0.373      & -0.239    \\
INSURANCE               & 0.018       & 0.017       & 0.012       & 0.018       & -0.002      & 0.001     \\
NBEDS                   & 0.136       & 0.131       & 0.217       & 0.089       & 0.080        & 0.109     \\
NBEDS/1000              & 0.138       & 0.121       & 0.209       & 0.092       & 0.069       & 0.088     \\
NHOSP                   & 0.074       & 0.119       & 0.165       & -0.009      & 0.028       & 0.021     \\
ESTBEDS                 & 0.136       & 0.131       & 0.217       & 0.089       & 0.080        & 0.109     \\
SENIOR                  & 0.104       & 0.028       & 0.033       & 0.152       & 0.164       & 0.156     \\
BLACK                   & 0.209       & 0.099       & 0.208       & 0.216       & 0.085       & 0.116     \\
HISPANICS               & -0.072      & -0.032      & -0.064      & -0.091      & -0.068      & -0.046    \\
POP\_DENSITY\_2010      & -0.302      & -0.210      & -0.175      & -0.272      & -0.231      & -0.216    \\
POVERTY                 & 0.128       & 0.059       & 0.200       & 0.148       & 0.031       & 0.080     \\
SVI\_MINORITY           & 0.029       & 0.069       & 0.081       & 0.010        & 0.006       & 0.087     \\
SVI\_OVERALL            & 0.140       & 0.158       & 0.238       & 0.125       & 0.113       & 0.179     \\
CASTHMA                 & 0.238       & 0.271       & 0.307       & 0.205       & 0.288       & 0.309     \\
HIGHCHOL                & 0.202       & 0.185       & 0.131       & 0.192       & 0.243       & 0.224     \\
DIABETES                & 0.200       & 0.200     & 0.178         & 0.206       & 0.232       & 0.230     \\
OBESITY                 & 0.199       & 0.205       & 0.194       & 0.191       & 0.229       & 0.212     \\
CANCER                  & -0.145      & -0.102      & -0.129      & -0.138      & -0.058      & -0.059    \\
STROKE                  & 0.123       & 0.125       & 0.117       & 0.167       & 0.143       & 0.135     \\
MHLTH                   & 0.110       & 0.161       & 0.186       & 0.076       & 0.175       & 0.222     \\
CSMOKING                & 0.189       & 0.258       & 0.249       & 0.160        & 0.289       & 0.323     \\
CHOLSCREEN              & 0.164       & 0.001       & -0.026      & 0.187       & 0.070       & 0.038     \\
INSURANCE               & 0.018       & 0.017       & 0.012       & 0.018       & -0.002      & 0.001     \\
CHD                     & 0.225       & 0.284       & 0.233       & 0.178       & 0.327       & 0.337     \\
CHECKUP                 & 0.206       & 0.221       & 0.280       & 0.201       & 0.272       & 0.343     \\
KIDNEY                  & 0.045       & 0.028       & 0.032       & 0.022       & 0.071       & 0.053     \\
BINGE                   & 0.031       & 0.044       & 0.009       & 0.083       & 0.030       & 0.022     \\
LPA                     & 0.191       & 0.187       & 0.197       & 0.147       & 0.220       & 0.214     \\
ARTHRITIS               & 0.046       & 0.013       & -0.021      & 0.061       & 0.089       & 0.079     \\
BPMED                   & 0.060       & 0.080        & 0.124       & 0.056       & 0.073      & 0.090      \\
PHLTH                   & 0.064       & 0.132       & 0.104       & 0.022       & 0.129       & 0.114     \\
BPHIGH                  & 0.056       & 0.051       & 0.048       & 0.058       & 0.084       & 0.068     \\
COPD                    & 0.163       & 0.163       & 0.153       & 0.175       & 0.220       & 0.240     \\
NORM\_POP               & -0.305      & -0.258      & -0.218      & -0.193      & -0.218      & -0.099  \\
\bottomrule
\end{tabular}
}
\end{table}

>>>>

\usepackage[table]{xcolor}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{rotating, multirow}
\usepackage{colortbl}

@@

\input{response-v3-content}

@@

To establish the feasibility of using deep learning, we have, in our
analysis, verified that through deep learning we can recreate the
predictions identified by a model-based approach using empirical
fits. In this comprehensive analysis, we obtain predictions through
deep learning while using data generated from epidemiological models
as discussed in \cite{marsland20-covid-paper}. The time series we used
was 100 days long, and a multi-layered Long Short-Term Memory (LSTM)
recurrent network was used. Our prediction approach differs by
learning not only from the demographics (fixed data for each city) and
time-dependent data but by integrating the population model for the
underlying prediction, as shown in Figure \ref{fig:magic-1}. Such model
predictive integration capability is important in any application with
multiple time scales. For example, this allows us to integrate
multi-scale time effects into the forecast, which could address the
combination of general forecasts and the next time steps of choice,
such as days, weeks, months, or longer. For this analysis, we used 37
of the 110 cities with reliable empirical fits (that we derived from
an epidemiological model and not deep learning
\cite{marsland20-covid-paper}) for the cases and deaths data up to
April 15, 2020. Our analysis identified a sophisticated single deep
learning time evolution operator that can describe these 37 separate
data sets, and smooth fitted data leads to very accurate deep learning
descriptions via LSTM as depicted in both Figure \ref{fig:magic-1}
that are smaller than 1\% from the model predictions.

@@

\begin{table}[!p]
\caption{Best models for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey.}
\label{tab:top-1}
\bigskip
\centering
%\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{Place/\\Rank} & 
\makecell{Risk Factor} &  
\makecell{RMSE\\Cases} &  
 \makecell{RMSE\\Death} &  
 \makecell[r]{Cumu-\\lative \\Error \\Cases} &  
 \makecell[r]{Cumu-\\lative \\Error \\Death} &  
 \makecell[r]{Number of \\Days as \\Input} \\
\midrule
1 & POP\_DENSITY\_2010 & 0.055443 & 0.067617 & 7.94 & 22.74 & 5 \\
2 & PHLTH & 0.054795 & 0.068529 & 8.01 & 20.55 & 3 \\
3 & INSURANCE & 0.055199 & 0.067641 & 8.07 & 20.43 & 4 \\
4 & PERCENTBLACKS & 0.055028 & 0.067207 & 8.12 & 20.17 & 4 \\
5 & PERCENTHISPANICS & 0.056061 & 0.067489 & 8.22 & 19.04 & 5 \\
6 & NHOSP & 0.054621 & 0.068255 & 8.26 & 20.10 & 3 \\
7 & INSURANCE & 0.054557 & 0.068572 & 8.34 & 24.61 & 3 \\
8 & CHOLSCREEN & 0.054712 & 0.069121 & 8.37 & 21.70 & 3 \\
9 & DIABETES & 0.054586 & 0.067636 & 8.40 & 20.50 & 5 \\
10 & STROKE & 0.054160 & 0.066906 & 8.45 & 20.16 & 4 \\
11 & CHD & 0.054027 & 0.067642 & 8.46 & 21.93 & 5 \\
12 & CHECKUP & 0.054927 & 0.067652 & 8.51 & 20.42 & 5 \\
13 & NBEDS & 0.054146 & 0.067536 & 8.54 & 21.15 & 5 \\
14 & SVI\_OVERALL & 0.054795 & 0.067676 & 8.58 & 19.94 & 5 \\
15 & KIDNEY & 0.054600 & 0.067565 & 8.61 & 21.86 & 5 \\
16 & CASTHMA & 0.055089 & 0.068248 & 8.63 & 22.58 & 5 \\
17 & BLACK & 0.054547 & 0.068266 & 8.68 & 20.49 & 3 \\
18 & BPMED & 0.054685 & 0.068920 & 8.74 & 21.69 & 3 \\
19 & LPA & 0.054323 & 0.067149 & 8.75 & 19.54 & 5 \\
20 & CSMOKING & 0.053716 & 0.066898 & 8.75 & 19.01 & 5 \\
21 & ARTHRITIS & 0.054053 & 0.068716 & 8.83 & 20.94 & 3 \\
22 & POVERTY & 0.055029 & 0.068672 & 8.84 & 22.02 & 3 \\
23 & ESTBEDS & 0.054545 & 0.067267 & 8.86 & 21.22 & 4 \\
24 & MHLTH & 0.054864 & 0.068742 & 8.90 & 19.53 & 3 \\
25 & SENIOR & 0.055079 & 0.069024 & 8.93 & 21.23 & 3 \\
26 & COPD & 0.055589 & 0.068250 & 8.93 & 23.26 & 4 \\
27 & CANCER & 0.054500 & 0.068363 & 8.99 & 19.70 & 3 \\
28 & NBEDS/1000 & 0.054397 & 0.068491 & 9.05 & 21.89 & 3 \\
29 & BPHIGH & 0.054609 & 0.067346 & 9.13 & 21.29 & 4 \\
30 & NONE &\rowcolor{black!5} 0.054868 & 0.069432 & 9.26 & 25.88 & 3 \\
31 & HIGHCHOL & 0.056429 & 0.069345 & 9.28 & 20.98 & 3 \\
32 & BINGE & 0.054317 & 0.067638 & 9.32 & 21.00 & 5 \\
33 & OBESITY & 0.054469 & 0.068540 & 9.50 & 20.65 & 3 \\
34 & SVI\_MINORITY & 0.055094 & 0.067443 & 9.54 & 21.53 & 4 \\
\bottomrule
\end{tabular}
%}
\end{table}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{images/predict/Length_5-fields_pop_density_2010-Cumulative-Cases.png}
    %\vspace{-1cm}
    \caption{Model prediction error for cumulative cases when
      including one risk factor - POP\_DENSITY\_2010 in the features,
      in addition to the past number of cases and deaths.}
    \label{fig:error-case-popdensity}
\end{figure}

\begin{figure}[!h]    
    \centering
    \includegraphics[width=0.5\textwidth]{images/predict/Length_5-fields_pop_density_2010-Cumulative-Death.png}
    %\vspace{-1cm}
    \caption{Model prediction error for cumulative deaths when
      including one risk factor - POP\_DENSITY\_2010 in the features,
      in addition to the past number of cases and deaths.}
    \label{fig:error-death-popdensity}
\end{figure}

@@

In our next comprehensive analysis, we try to answer Question
\ref{q:4} if all risk factors impact the analysis. Thus we feed the
covariates into the recurrent network as additional features. In
addition to cases and deaths, this leads to 35 variables, including 33
unique covariates. We looked at two related ways of doing this: (a)
have covariates fixed in time (static), and (b) multiply the
covariates by the times series (we used the average of cases and
deaths)

@@

\begin{table}[!h]
\caption{Correlation Matrix}
\bigskip
\label{tab:corr-matrix}
\begin{center}
\begin{tabular}{lrrrrrr}
\toprule
Risk Factor		&
\makecell{Past\\Case} &
\makecell{Now\\Case} &
\makecell{Future\\Case} &
\makecell{Past\\Death} &
\makecell{Now\\Death} &
\makecell{Future\\Death} \\
\midrule
PVI                     & -0.425      & -0.292      & -0.124      & -0.434      & -0.373      & -0.239    \\
INSURANCE               & 0.018       & 0.017       & 0.012       & 0.018       & -0.002      & 0.001     \\
NBEDS                   & 0.136       & 0.131       & 0.217       & 0.089       & 0.080        & 0.109     \\
NBEDS/1000              & 0.138       & 0.121       & 0.209       & 0.092       & 0.069       & 0.088     \\
NHOSP                   & 0.074       & 0.119       & 0.165       & -0.009      & 0.028       & 0.021     \\
ESTBEDS                 & 0.136       & 0.131       & 0.217       & 0.089       & 0.080        & 0.109     \\
SENIOR                  & 0.104       & 0.028       & 0.033       & 0.152       & 0.164       & 0.156     \\
BLACK                   & 0.209       & 0.099       & 0.208       & 0.216       & 0.085       & 0.116     \\
HISPANICS               & -0.072      & -0.032      & -0.064      & -0.091      & -0.068      & -0.046    \\
POP\_DENSITY\_2010      & -0.302      & -0.210      & -0.175      & -0.272      & -0.231      & -0.216    \\
POVERTY                 & 0.128       & 0.059       & 0.200       & 0.148       & 0.031       & 0.080     \\
SVI\_MINORITY           & 0.029       & 0.069       & 0.081       & 0.010        & 0.006       & 0.087     \\
SVI\_OVERALL            & 0.140       & 0.158       & 0.238       & 0.125       & 0.113       & 0.179     \\
CASTHMA                 & 0.238       & 0.271       & 0.307       & 0.205       & 0.288       & 0.309     \\
HIGHCHOL                & 0.202       & 0.185       & 0.131       & 0.192       & 0.243       & 0.224     \\
DIABETES                & 0.200       & 0.200     & 0.178         & 0.206       & 0.232       & 0.230     \\
OBESITY                 & 0.199       & 0.205       & 0.194       & 0.191       & 0.229       & 0.212     \\
CANCER                  & -0.145      & -0.102      & -0.129      & -0.138      & -0.058      & -0.059    \\
STROKE                  & 0.123       & 0.125       & 0.117       & 0.167       & 0.143       & 0.135     \\
MHLTH                   & 0.110       & 0.161       & 0.186       & 0.076       & 0.175       & 0.222     \\
CSMOKING                & 0.189       & 0.258       & 0.249       & 0.160        & 0.289       & 0.323     \\
CHOLSCREEN              & 0.164       & 0.001       & -0.026      & 0.187       & 0.070       & 0.038     \\
INSURANCE               & 0.018       & 0.017       & 0.012       & 0.018       & -0.002      & 0.001     \\
CHD                     & 0.225       & 0.284       & 0.233       & 0.178       & 0.327       & 0.337     \\
CHECKUP                 & 0.206       & 0.221       & 0.280       & 0.201       & 0.272       & 0.343     \\
KIDNEY                  & 0.045       & 0.028       & 0.032       & 0.022       & 0.071       & 0.053     \\
BINGE                   & 0.031       & 0.044       & 0.009       & 0.083       & 0.030       & 0.022     \\
LPA                     & 0.191       & 0.187       & 0.197       & 0.147       & 0.220       & 0.214     \\
ARTHRITIS               & 0.046       & 0.013       & -0.021      & 0.061       & 0.089       & 0.079     \\
BPMED                   & 0.060       & 0.080        & 0.124       & 0.056       & 0.073      & 0.090      \\
PHLTH                   & 0.064       & 0.132       & 0.104       & 0.022       & 0.129       & 0.114     \\
BPHIGH                  & 0.056       & 0.051       & 0.048       & 0.058       & 0.084       & 0.068     \\
COPD                    & 0.163       & 0.163       & 0.153       & 0.175       & 0.220       & 0.240     \\
NORM\_POP               & -0.305      & -0.258      & -0.218      & -0.193      & -0.218      & -0.099  \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
